### 2012-07-17 master-thesis/thesis/sections/cap-conclusions.tex

\chapter{Conclusions}
\label{cap:conclusions}

%---------------------------------------------%