### 2012-08-09 articles/CAMAD2102/article-version2.tex
More fixing and improving

%% bare_conf.tex
%% V1.3
%% 2007/01/11
%% by Michael Shell
%% See:
% Also note that the "draftcls" or "draftclsnofoot", not "draft", option
% should be used if it is desired that the figures are to be displayed in
% draft mode.
%
\documentclass[conference]{IEEEtran}

% *** GRAPHICS RELATED PACKAGES ***
%
\ifCLASSINFOpdf
\else

\fi

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

%%  my packages  %%
%\usepackage{sbc-template}

\usepackage{graphicx,url}

%\usepackage[brazil]{babel}
\usepackage[english]{babel}
\usepackage[latin1]{inputenc}
% *** SPECIALIZED LIST PACKAGES ***
%
\usepackage[lofdepth,lotdepth]{subfig}
\usepackage{amsthm} %%math
\usepackage{amsfonts}%%math
\newtheorem{mydef}{Definition}
%\newenvironment{definition}[1][Definition]{\begin{trivlist}
%\item[\hskip \labelsep {\bfseries #1}]}{\end{trivlist}}
\usepackage{comment}
\usepackage{multirow}
\usepackage{algorithm}
\usepackage{algorithmic}
%\floatname{algorithm}{Procedure}
\renewcommand{\algorithmicrequire}{\textbf{Input:}}
\renewcommand{\algorithmicensure}{\textbf{Output:}}

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
%\title{Análise de Desempenho de Coreografias de Serviços  no Modelo de Interação}
%\title{Performance Evaluation of Service Choreographies in the Interaction Model}
%\title{A Mechanism to Establish Network QoS Requirements of Service Choreographies}
%\title{A Methodology to Predict the Impact of Network QoS Attributes in Service Choreographies}
\title{ A Methodology to Define QoS and SLA Requirements in Service Choreographies }

% author names and affiliations
% use a multiple column layout for up to three different
% affiliations
\author{
  %\IEEEauthorblockN{Victoriano P. Diaz,  Daniel M. Batista}
  \IEEEauthorblockN{Victoriano P. Diaz and Daniel M. Batista}
  %\IEEEauthorblockN{Daniel M. Batista}
  \IEEEauthorblockA{Computer Science Department \\
  University of São Paulo (IME-USP)\\
  Postal Code 166281 -- Sao Paulo -- SP -- Brazil \\
  %Atlanta, Georgia 30332--0250\\
  Email:  \{alfonso7,batista\}@ime.usp.br
  }
  %\and
  %\IEEEauthorblockN{Daniel M. Batista}
  %\IEEEauthorblockA{Computer Science Department \\
  %University of São Paulo (IME-USP)\\
  %Postal Code 166281 -- Sao Paulo -- SP -- Brazil
  %Atlanta, Georgia 30332--0250\\
  %Email: http://www.michaelshell.org/contact.html
  %}
}
% conference papers do not typically use \thanks and this command
% is locked out in conference mode. If really needed, such as for
% the acknowledgment of grants, issue a \IEEEoverridecommandlockouts
% after \documentclass

% make the title area
\maketitle

\begin{abstract}
%\boldmath
Service choreography allows the composition of services in a collaborative way,  taking advantage of various benefits of P2P systems. The
 existing modeling approaches  to evaluate choreographies are the interconnection and interaction models. %Interaction models are  supported by BPMN standard version 2.0. 
However, these approaches don't evaluate choreographies focusing on QoS or in the earlier stages 
of development of the distributed application.
This paper proposes an approach to assess the impact of QoS attributes specified in a choreography interaction model. %using BPMN 2.0. For  this purpose, 
 %we use a representation of a choreography by a Generalized Stochastic Petri Net (GSPN). This representation
 %results from the composition of GSPN modules mapped from BPMN elements used in a choreography modeling. 
With our proposal it is possible to establish requirements for QoS and SLA in early stages of development in order to %reduce costs  and impact in later stages.
plan the capacity of the network elements connecting the hosts involved in the enactment and deployment of the choreography.
%, reducing the costs in later stages of the development.

\end{abstract}

\IEEEpeerreviewmaketitle

\section{Introduction} \label{sec:introduction}

The process to develop applications and complex systems evolved over time until converge to the approach know as
 Service Oriented Computing - SOC \cite{PapazoglouSOC2007a}. The service-based applications require compositions of services as a
 key factor to allow rapid development, loosely coupling and flexible integration with other systems. In this scenario, now,  a 
  single application is formed by several services spread over the network.

Among the various methods to compose services, Web services choreography is an efficient way to implement inter-organizational business
 processes, as the participants' business interactions are mutually independent (autonomous and heterogeneous) \cite{Telang2011}. A service
 choreography is a description of peer to peer interactions among existing services, i.e., in this model there isn't the role of 
 a central controller. The various services communicate with each other similarly to what occurs in a P2P application \cite{Barker2009}. 
%Business  processes are modeled and executed using workflow languages like BPEL for orchestration \cite{Barker2009}.

% Uma coreografia de serviços modela os processos visíveis publicamente e a troca de mensagens
% entre os participantes do ponto de vista global \cite{Wetzstein2010}.

 During the enactment of web services choreographies, the state of network elements (devices and links) plays a fundamental role. There must be guarantees 
 of Quality of Service - QoS so that there are advantages of using a decentralized business model. A common method to define
 guarantees between a service provider and a client (which may also be a service) is by means of a Service Level Agreement - SLA. After
 the choreography be specified,  constraints of QoS between each participant must be defined through SLAs \cite{Rosenberg2007}. To meet the SLA 
 agreements there must be some mechanism for management at runtime. This mechanism must involve monitoring, control and  
 decisions against violations or degradation of service quality. All the concern about guarantee of the QoS requirements of participants comes 
 from the fact that the QoS of composite service, represented by the choreography, depends directly on the QoS of the separate services.

 To reduce the number of SLA violations and the corrective measures to fulfill the QoS requirements during the enactment of choreographies, 
 the detection of non-functional failures in early stages of the development of the choreography can be employed. For example, 
  some non-functional failures (or undesirables properties)  are communication and network errors such as high latency and low bandwidth.
%, connection timed out, and so on .

 This paper presents a new approach to detect non-functional failures of the participants in a choreography during the design stage. Our approach
 conducts a performance evaluation by means of generalized stochastic Petri nets (GSPNs) as the intermediate representation of a choreography
 specified in BPMN2, a notational language to model business process, with the representation of the QoS model. Our work differs from 
 those found in the literature because it is employed in stages before the development (for example, in the modeling and design). With our 
 proposal it is possible to establish requirements for QoS and SLA in early stages in order to plan the capacity of the network connecting
 the hosts involved in the deployment of the application, reducing the costs and unexpected behaviors  in later stages of the development.

 This paper is organized as follows: Section~\ref{sec:basicConcepts}  introduces the basic concepts necessary to understand the
 proposal and presents the example of a choreography specified in BPMN2. Section~\ref{sec:relatedWorks} presents the related
 work. Section~\ref{sec:proposta} describes the proposal, which includes the formalization, the QoS model and the mapping for
 GSPN. Section~\ref{sec:PerformanceAnalysis} presents the application of the proposal in a scenario with the objective to evaluate this
 effectiveness. Finally, Section~\ref{sec:Conclusions} presents the conclusions and future work.

%\hfill mds

\section{Basic Concepts } \label{sec:basicConcepts}
  %\subsection{ Overview } %concepts
   %\subsection{ Conceitos } %concepts

  The two main approaches to compose services are choreographies and orchestrations. A service choreography allow the composition of services
 in a collaborative manner. A choreography describes the P2P interactions of the externally observable behavior of its participants. Different
 from orchestrations, choreographies don't have a single point of control or coordination \cite{Barker2009}. Choreographies tend to be more
 efficient than orchestrations because the probability to have a bottleneck in the system is reduced \cite{Barker2009b}.

  According to \cite{Kopp2011}, languages to specify choreographies can be classified based on two criteria: implementation and approach used
 in modeling. In terms of approach, there are languages with interaction models and languages with interconnection models. 
  To standardize the representation of choreographies, the BPMN graphical representation, currently in version 2.0, was proposed 
 (\url{http://www.bpmn.org/}). It supports the modeling of interconnection through collaboration, and the modeling of interaction through
 choreography of processes. In this paper it is assumed that the choreography is represented by models of interaction because they incur less
 redundancies and errors \cite{Decker2008}.
 %  Neste artigo será considerado que as coreografias são representadas por modelos de interação porque incorrem em menos redundâncias e erros  com uma modelagem
 %  mais simples que se encaixa adequadamente em metodologias de cima para baixo (\textit{top-down} ) \cite{Decker2008}.

  Figure~\ref{fig:ChoreographyElements} shows the BPMN elements considered in this paper. We consider only this subset because
  they are enough to evaluate our proposal. To the reader interested, in \cite{OMG2011} all the elements allowed in BPMN choreographies are presented.

  \begin{figure}[!h]
      \centering
      %\includegraphics[width=.90\textwidth]{Choreography_CaseStudy.png}
      \includegraphics[width=0.5\textwidth]{./figures/BPMNBasicChoroegraphy.png}
      %\caption{Elementos BPMN para modelar coreografias de processos}
      \caption{BPMN elements for modeling  choreographies }
      \label{fig:ChoreographyElements}
  \end{figure}

Figure~\ref{fig:choreographyExample} presents an example of choreography modeled with elements presented in Figure~\ref{fig:ChoreographyElements}. It 
 is a choreography for the offer of investment in stock market presented in \cite{Koppa}. First, a 
 financial advisor offers an investment to a customer and, after a time interval, the customer decides to accept or reject such a proposal, thereby ending the process.

 In the scenario of Figure~\ref{fig:choreographyExample}, basic elements of choreography processes are used, such as three choreography
 tasks, two events (\emph{start} and \emph{end}), six sequence flows and one gateway (\emph{Data-based XOR}). A choreography task 
represents an atomic interaction between two or more participants performed by exchanging messages via the network.

  \begin{figure}[!h]
      \centering
      %\includegraphics[width=.90\textwidth]{Choreography_CaseStudy.png}
      \includegraphics[width=0.5\textwidth]{./figures/BPMNChoreographyExample.png}
      %\caption{Modelagem BPMN para a coreografia de processo do cenário de exemplo}
      \caption{Example of choreography using BPMN2 elements.}
      \label{fig:choreographyExample}
  \end{figure}

%=====================%
\section{Related Works} \label{sec:relatedWorks}

 In \cite{Xia2009}, it is proposed an approach to predict  the QoS  of a Web service choreography specified in WSCI
 (\textit {Web Service Interaction Choreography}).  A GSPN is used as an intermediate model. An analytical evaluation is performed in the 
 QoS metrics tree and transformation rules to map activities are specified. A problem with the proposal presented in \cite{Xia2009}
  is that the WSCI language has been discontinued.

In \cite{Xia2011} it is  proposed an approach to predict the reliability of composite services specified in WS-CDL. It is also
 presented a complete translation of a choreography specified in WS-CDL to a GSPN. Similar to \cite{Xia2011}, we also use the formalism
 of Petri nets as an intermediate representation of a choreography. Different from \cite{Xia2011}, our proposal is focused in 
 QoS attributes of communication network and we use the interaction model of BPMN2 as choreography language.  
%differ from \cite{Xia2011} in terms of the objective. We define QoS requirements, while

\cite{Bruneo} and \cite{Valero2012} present a proposal to evaluate the performance of choreographies represented in WS-CDL using Petri nets. However,
 WS-CDL is a limited language since it is more oriented to implementation, instead interaction between the participants. Our proposal is 
 oriented to BPMN. For the best of our knowledge, there is no proposal that evaluate the performance of choreographies specified in BPMN.

%=================================%
\section{ Methodology to evaluate the performance of service choreographies} \label{sec:proposta}

The contribution of this paper is  a new methodology to evaluate the performance of choreographies with the objective of
 define  QoS requirements. The evaluation is performed in the design stage of the choreography by using GSPNs. The proposal is based 
on the execution of the following steps:

\begin{enumerate}

  \item Mapping of the choreography specified in BPMN 2.0 to a GSPN. The GSPN includes the intermediate representation of the choreography
 and the QoS model that takes into account time and communication constraints of the interactions among participants (each participant 
 represents a peer that is active during the enactment of the choreography).

  \item Definition of weights in the transitions with time constraints in the GSPN. In this step, the probability distributions and weights
 in the interactions related to communications and message exchange are defined.

 \item Simulations of scenarios, one with a minimum configuration of failures and another with a higher possibility of failure. Simulations
  are used because choreographies in the real world represent complex processes, which generate complex petri nets with many states 
  that become in difficult analytical evaluations \cite{Oliveira2009}.

%The simulations are employed to retrieve approximated performance measures with certain confidence level
\end{enumerate}

The next subsections present information needed to the execution of all the steps. Subsection~\ref{subsec:formalization} defines
 choreographies of processes in BPMN. Subsection~\ref{subsec:QoSModel} describes the QoS model  considered in our
 proposal, and  Subsection~\ref{subsec:mapping} presents the algorithm executed to map BPMN elements to a GSPN.

 %\subsection{ Formalization }   \label{subsec:formalization} %%or BPMN Chor
\subsection{ Definition of choreographies specified in BPMN }   \label{subsec:formalization}

 Based on Figure 1 and on the work presented in \cite{Dijkman2007}, choreographies of BPMN processes can be defined as explained below:

%. Para tanto, a seguir vamos definir o \textit{core process choreography BPMN} (baseado em \cite{Dijkman2007}).

    \begin{mydef}
      A choreography of BPMN processes is a tuple: \\
       $PC = (\mathcal{O, A, E, G, T}, \{e^S\}, \mathcal{E}^I, \{e^E\}, \mathcal{E}^{I_M}, \mathcal{E}^{I_T}, \mathcal{G}^F, \mathcal{G}^J,$
       $\mathcal{G}^X, \mathcal{G}^M, \mathcal{G}^V, \mathcal{F} )$ where:
      \begin{itemize}
	\item $\mathcal{O}$ is the set of objects, which is partitioned in three disjoint sets: activities ($\mathcal{A}$), events $\mathcal{E}$
	and \emph{gateways} $\mathcal{G}$.

	\item $\mathcal{A}$, is the set of \textit{Choreography Tasks} ($\mathcal{T}$), \textit{Sub-choreographies} and \textit{Call Choreographies}. However, 
	\textit{Sub-choreographies} and \textit{Call Choreographies} aren't taken into account in this work, then $\mathcal{A}$ it is the same as $\mathcal{T}$.

	\item $\mathcal{E}$ is the set of events. It is partitioned in three disjoint sets: Start event {$e^\mathcal{S}$}, Intermediate
	events $\mathcal{E^I}$ and End event {$e^\mathcal{E}$}.

	\item $\mathcal{E^I}$ is partitioned in the disjoint sets of events of the type messages $\mathcal{E}^{I_M}$ and timer $\mathcal{E}^{I_T}$.

	\item $\mathcal{G}$ é the set of gateways and it is partitioned in the 	five disjoints sets of parallel fork gateways $\mathcal{G}^F$,
	parallel join gateways $\mathcal{G}^J$, data-based XOR gateways $\mathcal{G}^X$, XOR merge gateways $\mathcal{G}^V$ and event-based
	XOR gateways $\mathcal{G}^M$.
	\item $\mathcal{F} \subseteq \mathcal{O}x\mathcal{O}$ is the control flow relation, i.e. a set of sequence flows connecting objects,
      \end{itemize}
    \end{mydef}

%Uma coreografia de processos é um grafo orientado com nós (objetos) $ \mathcal{O}$ e arestas (fluxos de sequencias) $\mathcal{F}$. para qualquer
% nó $x \in \mathcal{O}$, os nós de entrada de $x$ são dados por $in(x) = \{ y \in \mathcal{O} | y\mathcal{F}x\}$ e os nós de saída são dados por
% $out(x) = \{ y \in \mathcal{O} | x\mathcal{F}y\}$.

%%   GSPN Definition
%    \begin{mydef}
%    \end{mydef}

 %    Nesta definição,  não estão incluídas tarefas de orquestrações (tarefas tradicionais do BPMN), alguns eventos e nem menos exceções,  já que não
 %   são parte de uma coreografia de processos. %% Mehorar

\subsection{QoS model } \label{subsec:QoSModel}

%Descrição do modelo de QoS e os atributos de QoS levados em consideração...

%O validade do formato da mensagem recebida é considerada no segundo aspecto e no último aspecto, somente o tempo de execução  da operação do serviço é considerado.
%A tabela \ref{tableQoS} resume o modelo de QoS descrito e que será utilizado no mapeamento para a GSPN.

 In order to analyze the non-functional behavior of choreographies, it is necessary to define the QoS attributes considered. In this
 paper we consider three entities with QoS attributes: service, network and message. In the operation of the service, the QoS attribute
 considered is the  time needed to complete the service.  In the network, the QoS attributes considered are delay and errors in communication, 
since both significantly  affect the response time of the choreography. In the message, the QoS attribute considered is the format of the message, 
 since the message need to be validated.

\subsection{ Mapping of BPMN choreography to GSPN} \label{subsec:mapping}

 Figure~\ref{fig:MappingEventsAndGateways} presents the mapping of gateways and events (choreography elements) to the respective modules of Petri nets.
Figure~\ref{fig:MappingTaskChoreographiesQoS1} exemplifies the mapping of two different choreography tasks (choreography tasks are the basic
building blocks of choreographies since they generate messages and service invocation). In Figure~\ref{fig:MappingTaskChoreographiesQoS1}, positions $A$,
$A_2$ and $B$ represent the participants. Positions $C_{init}$, $C_{end}$, $C_{init2}$, $C_{end2}$, $C_{endOk}$ and $C_{endOk2}$ represent the start, end and the
correct validation of the communication between two participants. Transitions of type $T_{send}$ and $T_{receive}$ represent the actions of sending and
receiving messages. Transitions $T_{msg}$ represent the transmission of the message by the communication channel.

% Esses módulos de redes de Petri devem ser compostos de acordo à sequência de fluxo e os \textit{gateways} presentes na
% modelagem da coreografia tal como está ilustrada na Figura \ref{fig:MappingEventsAndGateways}. Uma rede de Petri composta representa uma
% coreografia de processos completa, e como um processo deve ter uma posição de começo e uma de fim.

  \begin{figure}[!h]
      \centering
      %\includegraphics[width=.90\textwidth]{Choreography_CaseStudy.png}
      \includegraphics[width=0.5\textwidth]{./figures/BPMNChoreographies2_2.png}
      \caption{Mapping of events and gateways elements to modules of Petri nets}
      %\caption{Mapeamento de eventos e \emph{gateways} para módulos de redes de Petri}
      \label{fig:MappingEventsAndGateways}
  \end{figure}

  \begin{figure}[!h]
      \centering
      %\includegraphics[width=.90\textwidth]{Choreography_CaseStudy.png}
      \includegraphics[width=0.5\textwidth]{./figures/BPMNChoreographies1-QoS.png}
      %\caption{GSPN modules of task choreograpies including the QoS model}
      \caption{  Mapping of two different choreography tasks with the QoS model}
      \label{fig:MappingTaskChoreographiesQoS1}
  \end{figure}

 The Step 1 of our methodology is executed by the Algorithm~\ref{alg2}. The algorithm maps a choreography of processes specified in BPMN 2.0 
to a GSPN including the QoS model. Each BPMN element is mapped to its respective module of Petri nets and the modules are composed as a
function of the sequence flows and the gateways. Finally, the Start and the End events are included.

%Basicamente, a GSPN da coreografia toda é obtida mediante a composição dos módulos de GSPN.

     %\textbf{Output:} Rede de Petri Estocástica Generalizada $GSPN$. \\
    \begin{algorithm}[!h]
      %\caption{Mapping a Interconnection Choreography in BPMN onto a GSPN with QoS model}
      %\caption{ { \small Mapeamento de uma coreografia especificada em BPMN 2.0 para uma GSPN com o modelo de QoS } }
      \caption{ { \small Mapping of choreography specified in BPMN 2.0 to a GSPN with QoS model} }
      \label{alg2}
      \begin{algorithmic}
      { \small
	%\REQUIRE Process Choreography $\mathcal{PC}$ in BPMN 2.0.
	\REQUIRE Process Choreography $PC = (\mathcal{O, A, E, G, T, S}, \{e^S\}, \mathcal{E}^I,$ $\{e^E\}, \mathcal{E}^{I_M}, \mathcal{E}^{I_T},
         \mathcal{G}^F, \mathcal{G}^J,$ $\mathcal{G}^X, \mathcal{G}^M, \mathcal{G}^V, \mathcal{F} )$ in BPMN 2.0.

	\ENSURE Generalized Stochastic Petri Net $GSPN_{QoS}$.

	\STATE Consider $CT_i \in \mathcal{T} $, $G_j \in \mathcal{G} $ and $E_k \in \mathcal{E}$. $i, j, k \in \mathbb{N} $
	%\STATE Let $PNQoS(CT_i)$ a respective GSPN including QoS from type of $CT_i$ and according to mapping rules (Figure \ref{fig:MappingTaskChoreographiesQoS1}).
	%\STATE Let $PNQoS(CT_i)$ be a respective GSPN including QoS from type $CT_i$ and according to mapping rules (Figure \ref{fig:MappingTaskChoreographiesQoS1}).
	\STATE Consider $PNQoS(CT_i)$ the respective GSPN including QoS as a function of the type of $CT_i$ and the mapping rules specified on Figure \ref{fig:MappingTaskChoreographiesQoS1}.
	\STATE Consider $PNQoS(G_j)$ the respective GSPN including QoS as a function of the type of $G_j$ and the mapping rules specified in Figure \ref{fig:MappingEventsAndGateways}.
	\STATE Consider $PNQoS(E_k)$ the respective GSPN including QoS as a function of the type of $E_k$ and the mapping rules specified in Figure \ref{fig:MappingEventsAndGateways}.
	\STATE Consider $\oplus$ The binary operator of composition of two GSPNs that returns other GSPN.

	\STATE $GSPN_{QoS} \leftarrow  Empty\  Petri\  Net$

	%\FOR { $CT_i \in T $ }
	 % \STATE $GSPN_{QoS} \leftarrow GSPN_{QoS} \oplus PNQoS(CT_i)$
	%\ENDFOR

	\FOR { $CT_i \in \mathcal{T} $ }
	  \STATE $GSPN_{QoS} \leftarrow GSPN_{QoS} \oplus PNQoS(CT_i)$
          \STATE Add a arrival timed Transition at beginning of the $GSPN_{QoS}$.
	\ENDFOR

	\FOR{ $G_j \in \mathcal{G} $ }
          \STATE $GSPN_{QoS} \leftarrow GSPN_{QoS} \oplus PN(G_j)$
	\ENDFOR

	\FOR{ $E_k \in \mathcal{E} $ }
          \STATE $GSPN_{QoS} \leftarrow GSPN_{QoS} \oplus PN(E_k)$
	\ENDFOR

	\STATE Add a starting Place and immediate Transition at the beginning of the $GSPN_{QoS}$.
        \STATE Add a ending Place and immediate Transition at the end of the $GSPN_{QoS}$.

	\RETURN $GSPN_{QoS}$
      }
      \end{algorithmic}
    \end{algorithm}

%%%

%===========================================%
%\section{Analysis and Performance Evaluation}
\section{Performance Evaluation}  \label{sec:PerformanceAnalysis}

%Nesta seção são aplicados os passos descritos da nossa abordagem no cenário de exemplo da Seção \ref{serviceChoreography} .
To attest the efficacy of our proposal, in this section all the steps of our methodology are performed. The choreography
 presented in Figure~\ref{fig:choreographyExample} is used in the process  because that scenario is simple in order to show and understand better the
 proposed methodology, and includes at least one element of the three choreography  BPMN types considered (activities, gateways and events). Each one of the
 next subsections describes one of the three steps of  the proposal defined in the Section~\ref{sec:proposta}.

\subsection{ Mapping } \label{subsec:mapping_performance}

Figure~\ref{fig:ExampleChoreographyGSPNQoS} presents the GSPN obtained as the result of the mapping of the choreography of Figure~\ref{fig:choreographyExample}. Tables 
\ref{table:place_description} and \ref{table:transition_description} list the interpretation of all places and transitions of the GSPN. The places
 represent the participants, the start and the end of a communication channel in the interactions, the start and the end of the choreography and 
 the points of decision and errors according to the QoS model.

  \begin{figure}[!h]
      \centering
      %\includegraphics[width=.90\textwidth]{Choreography_CaseStudy.png}
      \includegraphics[width=0.5\textwidth]{./figures/BPMNChoreographyExample-QoS.png}
      %\caption{GSPN obtido da coreografia do cenário de exemplo}
      \caption{GSPN obtained from the choreography of Figure~\ref{fig:choreographyExample} }
      \label{fig:ExampleChoreographyGSPNQoS} 
  \end{figure}

\begin{table}[!h]
    \centering
    %\caption{Descrição das posições da GSPN do exemplo }
    \caption{  Description of the places in the GSPN of Figure~\ref{fig:ExampleChoreographyGSPNQoS} }
    \label{table:place_description}
    \begin{tabular}{|l|l|}
      \hline
       Place     &       Description \\
       \hline
       $Start$, $End$            &  Start and end of choreography process. \\
       $A_1$, $A_2$, $A_3$   &  Represents the participant ''Customer''. \\
       $B_1$, $B_2$, $B_3$   &  Represents the participant ''Financial Advisor''. \\
       $C_{i1}$, $C_{i2}$, $C_{i3}$   &  Start of the communication. \\
       $C_{f1}$, $C_{f2}$, $C_{f3}$   &  End of the communication. \\
       $C_{fok1}$, $C_{fok2}$, $C_{fok3}$   &  Valid message format . \\
       $C_{error1}$    & Communication error  between A and B. \\
       $C_{error2}$    & Communication error  between B and A. \\
       $C_{error3}$    & Communication error  between B and A. \\
       $M_{err1}$, $_2$, $_3$  & Invalid message format. \\
       $Xor$ e $Sync$  & \textit{Fork} and \textit{Join} of  exclusive gateway XOR. \\
      \hline
    \end{tabular}
\end{table}

\begin{table}[!h]
    \centering
    %\caption{Descrição de transições da GSPN do exemplo }
    \caption{  Description of the transitions in the GSPN of Figure~\ref{fig:ExampleChoreographyGSPNQoS} }
    \label{table:transition_description}
    \begin{tabular}{|l|l|}
      \hline
       Transition     &       Description \\
       \hline
       $T_{arrival1}$  &  Choreography instances arrival.\\
       $T_{arrival2}$, $_3$   & Choreography instances arrival and  Flow decisions.\\
       $T_{send1}$, $T_{send2}$ e $T_{send3}$   &  Activity of message sending. \\
       $T_{latency1}$, $_2$, $_3$   &     Network latency\\
       $T_{receive}$, $_2$, $_3$   &     Receive and service execution.\\	
       $T_{cerr1}$, $_2$, $_3$  &  Communication error. \\
       $T_{merr1}$, $_2$, $_3$ &  Format fault message. \\
       $T_{g}$   & Flow sequence continuation. \\
       $T_{ok1}, T_{ok2}, T_{ok3}$   &  Valid receive message . \\
       $T_{joining2}$, $_3$   &   Flow joining. \\
       %$T_{starting}$	 &  Starting choreography instance. \\
       $T_{ending}$      &  Ending choreography instance. \\

      \hline
    \end{tabular}
\end{table}

%\subsection{ Simulations } \label{subsec:simulations}
\subsection{ Definitions of weights } \label{subsec:simulations}

The next step is to define the weights (rates) representing the timed transitions and the priorities in the immediate transitions. These values
 depend on the knowledge about the environment where the choreography will be executed and the reliability of the independent
 services. In this paper we don't consider uncertainties in these values. Table~\ref{table:transitionsConfigurations} presents
 the definition of the weights and priorities in two scenarios, that will be needed to the simulations on the next step. Scenario 1 is configured
 to have less probability of failures than Scenario 2.

\begin{table}[!h]
    \centering
    %\caption{Configuração de pesos nos Cenários 1 e 2}
    \caption{Weights of Scenario 1 and Scenario 2}
    \label{table:transitionsConfigurations}
    \begin{tabular}{|l|l|l|}
      \cline{1-3}
		    &   \multicolumn{2}{|c|}{ Weights } \\
      \cline{1-3}
      %\hline
       \multicolumn{1}{|c|}{Transition }     		&      Scenario 1   &   Scenario 2 		\\
       \hline
       $T_{latency1}$, $T_{latency2}$, $T_{latency3}$ 	& 	0.99       &   0.94	\\
       $T_{cerr1}$, $T_{cerr2}$, $T_{cerr3}$      	&  	0.01 	   &   0.06	\\
       $T_{receive}$, $T_{receive2}$, $T_{receive3}$  &  	99  	   &    97	\\
       $T_{merr1}$, $T_{merr2}$, $T_{merr3}$   &   1 	   &    3	\\
       $T_{arrival2}$, $T_{arrival3}$   &  0.5  &	0.5	\\
      \hline
    \end{tabular}
\end{table}

\subsection{Simulations} \label{subsec:analysis}

To simulate the scenarios, 100 tokens are considered to each scenario at the place Start. Each token represents one instance of execution
 of the simulation. So, in each simulation, 100 concurrent instances of the scenarios were executed (multiple-server semantic). 
%Since  there are probabilities  involved in the experiments
The two scenarios were executed several times according to configurations in Table~\ref{table:transitionsConfigurations}. There were defined 1500
 fires and 10 replications. The Pipe2 \cite{Bonet2007} tool was used to model GSPNs and simulate the scenarios.

 Table~\ref{table:simulationsAll} shows the results of the simulations with confidence level of $95\%$. The numbers presented in 
the table show the average of the tokens that arrived at each one of the places of the GSPN.

%A tabela \ref{table:simulationsAll} mostra os principais resultados das simulações dos cenários 1 e 2 respectivamente nas quais se especificaram
% 1500 disparos e 10 replicações. Estes resultados mostram a média do número
%de \emph{tokens} que passaram em uma determinada posição e o intervalo de confiança para indicar a confiabilidade da estimativa.

\begin{table}[!h]
    \centering
    \caption{Results  (in \%) }
    \label{table:simulationsAll}
    \begin{tabular}{|l|l|l|l|l|}
	  \hline
		       & \multicolumn{2}{c|}{ Average number of tokens } & \multicolumn{2}{c|}{ 95\% Confidence interval  (+/-) }\\
	  \hline
	  Place      &  Scenario 1   &   Scenario 2  	 &  Scenario 1 	  &   Scenario 2 \\
	  \hline
	  $Start$ &    35,28	 &    40,15        &   5,83    	 &    6,23         \\
	  $End$   &    41,95	 &    38,78        &   2,53  	  &  3,82	\\
	  $A_1$   &    0,08	 &    0,08         &   0,00   	&  0,00     \\ %% após da execucao do Treceive 1
	  $B_2$	  &    0,04	&     0,01 		& 	0,01 &  0,01 \\ %% após da execucao do Treceive 2
	  $B_3$	  &    0,04	&     0,01  		& 0,01 	&  0,01 \\ %% após da execucao do Treceive 3
      \hline
	  $B_1$	  &    0,08	 &  3,23  		& 0,00	&  0,00	\\ %Primeiro participante em iniciar a interação
	  $A_2$	  &    0,04	 &  0,01		& 0,01	&  0,01	\\ %% aceito
	  $A_3$	  &    0,04	 &  0,01		& 0,01	&  0,01	\\ %%rejeitado
        \hline
	  $M_{err1}$	& 0,39	 & 0,91 		& 0,95	&  1,92	\\
	  $M_{err2}$	& 0,00	 & 0,93 		& 0,63		&  0,64	\\
	  $M_{err3}$	& 0,00   & 0,66 		& 0,87	&  0,74	\\
        \hline
	  $C_{err1}$	& 0,74 	 & 2,94 	& 0,82	&  2,02	\\
	  $C_{err2}$	& 0,00	 & 0,00 	& 0,67	&  1,75	\\
	  $C_{err3}$ 	& 0,78	 & 0.16 	& 0,92	&  1,52	\\
        \hline
	  $C_{i1}$ 	& 8,32	 & 8,90 	& 5,33	&  7,48	\\
      $C_{i2}$ 	& 0,63	 & 0,69 	& 0,23	&  0,52	\\
      $C_{i3}$ 	& 0,75	 & 8,90 	& 0,39	&  0,21	\\

	  \hline
	  \end{tabular}
\end{table}

%A tabela \ref{table:simulationsAll} mostra que no cenário $1$ em média  $1,51643\%$ ($Cerr1 + Cerr2 + Cerr3$) de instâncias de coreografias não terminaram o processo ($End$) por causa de erros na comunicação na rede entre as interações entre o participante Financeiro com o Cliente no momento de enviar a mensagem. Também, em média $0,39202\%$ ($Merr1 + Merr2 + Merr3$) de instâncias
%não terminaram o processo por causa de erros no formato de mensagem recebida.

By the results presented in Table~\ref{table:simulationsAll} it is possible to note that in the Scenario 1, an average of $1.52\%$ ($C_{err1} + C_{err2} + C_{err3}$) of 
the instances didn't  finish the process because of error in the network in the moment of the message exchanges. Also, an average
 of $0,39\%$ ($M_{err1} + M_{err2} + M_{err3}$) of the instances didn't finish the process because of error in the messages
 format, detected when the messages were received.

%Da mesma forma, no cenário $2$  houve uma perda de   $4,52167\%$ ($Cerr1 + Cerr2 + Cerr3$) e   $2,49416\%$ ($Merr1 + Merr2 + Merr3$)  de instâncias que não terminaram o processo por causa de erros de comunicação e de formato de mensagem respectivamente. Resultando que efetivamente houve uma maior perda de instâncias de coreografias do que no cenário $1$. No entanto, o que se manteve foi  o gargalo achado no começo da comunicação na primeira interação ($C_{i1}$), retendo  em média $8,32316\%$ e $8,90167\%$ das instâncias antes de serem enviadas pela rede o que pode contribuir em um atraso do tempo de execução total do processo.
Similarly to the Scenario 1, in Scenario 2, an average of $3,10\%$ and $2,50\%$ of the instances didn't finish the process because of error in 
the network and in the message format respectively. As expected, the results are inline with the parameters
 defined to each scenario (Table~\ref{table:transitionsConfigurations}).

One important finding with the simulation of these scenarios, and that can be useful to developers and network administrators is related with the existing 
bottleneck in the first iteration ($C_{i1}$), which avoided $8,32\%$ and $8,90\%$ of the instances for being send via
 network. This information can affect the implementation of the definite choreography or even define new QoS policies to be
 configured in the network to reduce the delay of the executions.

%Então, como pode se apreciar dá para reconhecer requerimentos de QoS a partir da análise da GSPN obtida. Por exemplo, a resposta de aceitação de uma oferta de
%A partir dos resultados obtidos das simulações sobre a GSPN que representa a coreografia e o modelo de QoS, dá para reconhecer alguns requerimentos de QoS. Por exemplo, a resposta de aceitação de uma oferta por um Cliente é critica para conseguir lucro na empresa, é por isso que  as falhas devem ser  mínimas e podem ser garantidas mediante o estabelecimento de restrições de QoS e SLA dos serviços entre os participantes.

%====================================%
%\section{Conclusões e Trabalhos Futuros}
\section{Conclusions and Future Works} \label{sec:Conclusions}
  This paper presented a new methodology to evaluate the impact of QoS attributes in choreographies of process specified in BPMN 2.0 and
following the interaction model. This methodology is executed in the earlier stages of the development with the objective of defining
QoS constraints and SLAs.  This definition allows a reduction in the failures when the choreography will be deployed and enacted. The
methodology uses a new algorithm to map the choreography into a GSPN. The methodology was employed in two scenarios and the results
confirmed its efficacy, since it was observed the presence of a bottleneck in the choreography that can be solved by changing the
interaction in the development or by employing QoS mechanisms in the network to prioritize the traffic affected by the bottleneck.

 As future work we will expand our methodology to support generic  probability distributions in the decision points.

%Um  trabalho futuro que já se encontra em desenvolvimento pelos autores é considerar aqueles elementos BPMN 2.0 para especificar coreografias de processos
%que não foram considerados, tais como \emph{sub-choreographies}, \emph{call choreographies}, laços, os outros tipos de \emph{task choreography},
%entre outros. Além do mais, precisa-se outro tipo de rede de petri para tratar o problema de correlação de instâncias de coreografias e suporte
% de outros tipos de distribuição de probabilidade nas transições com tempo.  Nessa linha, o uso de redes de petri coloridas pode ajudar
% a resolver esses problemas.
%Finalmente, precisamos realizar uma análise mais minuciosa da GSPN obtida e  aplicar nossa proposta para modelos de interconexão de coreografias  e assim realizar uma comparação.

\section*{Acknowledgment}
This work is supported by HP Brasil under technical cooperation agreement number HP-045/12 (Baile Project - \url{http://ccsl.ime.usp.br/baile/}) 
and the European Community's Seventh Framework Programme FP7/2007-2013 under grant agreement number 257178 (Project CHOReOS - Large Scale 
Choreographies for the Future Internet - \url{http://www.choreos.eu}).

% This work is supported by the European Community's Seventh Framework Programme FP7/2007-2013 under grant agreement number
%257178 (project CHOReOS - Large Scale Choreographies for the Future Internet - \url{www.choreos.eu}).

%The authors would like to thank the CHOReOS Project of the  FP7 European program FP7-ICT-2009-5 for financial support. We  would also like to thank to
%staff of CHOReOS USP partner.

\bibliographystyle{IEEEtran}
%\bibliographystyle{acm}
%\bibliography{sbc-template}
\bibliography{references5.bib}

% that's all folks
\end{document}